---
title: "Assignment 1"
author: "Dana El-khatib"
date: "10-10-2023"
output: html_document
---

```{r setup, include=FALSE} 
# this chunk contains code that sets global options for the entire .Rmd. 
# we use include=FALSE to suppress it from the top of the document, but it will still appear in the appendix. 

knitr::opts_chunk$set(echo = FALSE) # actually set the global chunk options. 
# we set echo=FALSE to suppress code such that it by default does not appear throughout the document. 
# note: this is different from .Rmd default
```

Exercise 1:

My forked version of the assignment template repository can be found at https://github.com/delkh/472_assignment_template.git

Excercise 2: 
```{r}
suppressMessages(library(tidyverse))
data('Orange')
?head()
head(Orange)
```

The data in Orange is tidy data, because it satisfies the following 3 condition:
1.Each variable must have its own column - age and circumference are two separate columns
2. Each observation must have its own row - for each number tree, the age and corresponding circumference is a seperate row.
3. Each value must have its own cell - each unique number (age or circumference) is in it's own cell

```{r}
data2 <- Orange %>%
  pivot_longer(c(age,circumference ), names_to = "measurement", values_to = "values")

head(data2)
```

The new version of the data (data2) is now untidy because two different types of values (age and circumference) are included in the same column. Additionally, two different variables (age and circumference) are in the same column (measurement)



Exercise 3:


```{r, echo = TRUE}

data('ToothGrowth') #loading a new dataset names ToothGrowth
# note that prior to using the summarize function, I tested that there are an equal number of observations per dose for each type of supplement using the following code  #count(new_tbl,dose =0.5) 
#count(new_tbl,dose =1)
#count(new_tbl,dose =2)
new_tbl <- ToothGrowth %>% group_by(supp) %>% summarise(avg_tooth_growth=mean(len))#grouping data in ToothGrowth by supplement type and taking the average of the length growth of each supplement. 

head(new_tbl) #visualizing the outcome

```

The above simple manipulation leads to the conclusion that OJ supplements on average has a higher impact on tooth growth than VC supplements, which was not immediately obvious in the raw data. 

## Appendix: All code in this assignment

```{r}
suppressMessages(library(tidyverse)) #loading tidyverse

# exercise 2
data('Orange') #loading r data set
head(Orange) #visualizing Orange dataset 


?pivot_longer #to understand pivot_longer
data2 <- Orange %>%
  pivot_longer(c(age,circumference ), names_to = "measurement", values_to = "values") #making the data untidy using pivot_longer

head(data2) #visualizing the new untidy data




```

